Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-03-10 paper.qmd
Updated the some code with new functions

#umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_15_min_dist_0.99.rds")

@@

tsne_two_curvy_scaled <- two_curvy_scaled_obj$scaled_nldr |>
  mutate(ID = row_number())
lim1 <- two_curvy_scaled_obj$lim1
lim2 <- two_curvy_scaled_obj$lim2
r2 <- diff(lim2)/diff(lim1)

@@

trimesh_two_curvy_removed1_umap_with_data <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000") +
  geom_point(data = tsne_two_curvy_scaled,
             aes(
               x = tSNE1,
               y = tSNE2
             ),
             color = "#636363",
             alpha = 0.5,
             size = 0.5) +
  #coord_equal() +
  # theme(
  #   aspect.ratio = 1
  #   ) +
  interior_annotation("b", sc_ltr_pos, cex = 1.5)
```

@@

df_bin_two_curvy_temp <- df_bin_two_curvy

### Compute highD distance
dist_vec <- proxy::dist(x = df_bin_two_curvy_temp[, -1], method = "Euclidean") |> as.vector()

from_vec <- c()
to_vec <- c()
num_obs <- 1:(NROW(df_bin_two_curvy_temp) - 1)

for (obs in num_obs) {

  from_val <- rep(obs, (NROW(df_bin_two_curvy_temp) - obs))
  if ((obs + 1) <= NROW(df_bin_two_curvy_temp)) {
    to_val <- (obs + 1):NROW(df_bin_two_curvy_temp)
  }
  from_vec <- append(from_vec, from_val)
  to_vec <- append(to_vec, to_val)

dist_highd <- tibble::tibble(from = from_vec, to = to_vec, dist_highd = dist_vec)

## To plot the distribution of distances

# Define bin width and starting point
start_point <- bin_width/2

# Create bins and calculate the mean value for each bin range
bins <- seq(start_point, max(distance_two_curvy$distance) + bin_width, 
            by = bin_width)
bin_labels <- bins[-length(bins)] + bin_width / 2  # mean of each bin range

distance_two_curvy_dist <- distance_two_curvy |> 
  mutate(bin = cut(distance, breaks = bins, include.lowest = TRUE, labels = bin_labels))

distance_two_curvy_dist <- left_join(distance_two_curvy_dist, dist_highd, by = c("from", "to"))

distance_points <- ggplot(
  distance_two_curvy_dist, 
  aes(x = distance,
      y = dist_highd)) +
  geom_point(alpha = 0.5) +
  geom_vline(xintercept = 2.5 * bin_width,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  geom_vline(xintercept = 2 * bin_width,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  ylab(expression(d^{(7)})) +
  xlab(expression(d^{(2)})) +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

@@

# hex_grid_coloured_two_curvy1 + 
#   hex_grid_coloured_two_curvy2 + 
#   hex_grid_coloured_two_curvy3 +
#   plot_layout(guides='collect', ncol = 3,
#               heights = c(2,1)) &
#   theme(legend.position='none', plot.tag = element_text(size = 8))

hex_grid_coloured_two_curvy2 +
  hex_grid_coloured_two_curvy1 +
  hex_grid_coloured_two_curvy3 +
  #boxplot1 + boxplot2 + boxplot3 +
  plot_layout(guides='collect', ncol = 3) &
  theme(legend.position='none', plot.tag = element_text(size = 12))
```

>>>>

@@

tsne_two_curvy_scaled <- two_curvy_scaled_obj$scaled_nldr |>
  mutate(ID = row_number())

lim1 <- two_curvy_scaled_obj$lim1
lim2 <- two_curvy_scaled_obj$lim2
r2 <- diff(lim2)/diff(lim1)

@@

trimesh_two_curvy_removed1_umap_with_data <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000") +
  geom_point(data = tsne_two_curvy_scaled,
             aes(
               x = tSNE1,
               y = tSNE2
             ),
             color = "#636363",
             alpha = 0.5,
             size = 0.5) +
  interior_annotation("b", sc_ltr_pos, cex = 1.5)
```

@@

df_bin_two_curvy_temp <- df_bin_two_curvy
```

@@

hex_grid_coloured_two_curvy2 +
  hex_grid_coloured_two_curvy1 +
  hex_grid_coloured_two_curvy3 +
  plot_layout(guides='collect', ncol = 3) &
  theme(legend.position='none', plot.tag = element_text(size = 12))
```
